The following set of analyses explore the article by Cheung (2008)

http://www.statmodel.com/download/MCheung.pdf

Cheung, M.W.L. (2008). A model for integrating fixed-, random-, and mixed-effects meta-analyses into structural equation modeling. Psychological Methods, 13, 182.

# Fixed effects meta-analysis
Let's create some data from the model for a fixed-effects meta-analysis.
```{r message=FALSE}
library(metafor)
```

```{r}
k <- 10
beta_fixed <- .2
meta_data <- data.frame(k=seq(k), sigma=runif(k, 0.05, 0.20))
meta_data$sigma_sq <- meta_data$sigma^2
meta_data$y <- rnorm(k, mean=beta_fixed, sd=meta_data$sigma)
```


```{r}
fixed_effects_meta_analysis <- function(y, sigma_sq) {
    # checks
    if (length(y) != length(sigma_sq)) {
        stop('length of y and sigma_sq should be equal')        
    }
    if (any(sigma_sq <= 0)) {
        stop('non-positive sigma_sq: All sigma_sq must be positive')        
    }
    
        
    k <- length(y)
    w <- 1/sigma_sq    
    beta_hat_fixed <- sum(w*y) / sum(w)
    s_squared_hat_fixed <- 1/ sum(w)
    s_hat_fixed <- sqrt(s_squared_hat_fixed)
    
    Z1 <- beta_hat_fixed / s_hat_fixed
    Z1_p_two_tail <- 1 - pnorm(abs(Z1))
    
    q <- w * (y - beta_hat_fixed)^2
    Q <- sum(q)
    Q_df <- k - 1
    Q_p <- 1 - pchisq(Q, df=Q_df)

    meta_table <- data.frame(y, sigma_sq, sigma=sqrt(sigma_sq), w, q)
    results <- list(meta_table=meta_table,
         beta_hat_fixed=beta_hat_fixed,
         s_squared_hat_fixed=s_squared_hat_fixed,
         Z1=c(Z1=Z1, p_two_tail=Z1_p_two_tail),
         Q=c(Q=Q, df=Q_df, p=Q_p))
    class(results) <- 'fixed_effects_meta_analysis'
    results
}

print.fixed_effects_meta_analysis <- function(x, digits=3, ...) {    
    print(lapply(x, function(X) round(X, digits=digits)))    
}

plot.fixed_effects_meta_analysis <- function(x, ...) {
    require(metafor)
    forest.default(x=x$meta_table$y,
                   se=x$meta_table$sigma,
                   refline=x$beta_hat_fixed,
                    ...)
}
                                          
meta1 <- fixed_effects_meta_analysis(y=meta_data$y, 
                           sigma_sq=meta_data$sigma_sq)
```

* The above code implements equations 2 to 5 in Cheung (2008).
* I created an S3 class. A `print` method is used to make display of the results a little easier to read. This is called by default when the return object printed to the console.
* A `plot` method produces a forest plot using the `forest.default` function in the `metafor` pakcage.

```{r}
plot(meta1)
meta1
```

# Models with covariates
```{r}
k <- 10
beta <- c(.2, .3)
meta_data_covariates <- data.frame(k=seq(k), 
                                   sigma=runif(k, 0.05, 0.20),
                                   x1=c(rep(1, k/2), rep(0, k/2)),
                                   x2=c(rep(0, k/2), rep(1, k/2)))
meta_data_covariates$sigma_sq <- meta_data$sigma^2
meta_data_covariates$beta <- as.matrix(meta_data_covariates[,c('x1', 'x2')]) %*% beta
meta_data_covariates$y <- rnorm(k, 
                                mean=meta_data_covariates$beta, 
                                sd=meta_data_covariates$sigma)
```

```{r}
fixed_effects_covariates_meta_analysis <- function(y, X, sigma_sq) {
    # y: vector of effect sizes of length k, 
    # X: design matrix (k rows by p columns)
    # Ve: vector of effect size variances of length k
    p <- ncol(X)
    k <- length(y)
    
    Ve <- diag(sigma_sq)
    beta_hat <- solve(t(X) %*% solve(Ve) %*% X) %*% t(X) %*% solve(Ve) %*% y

    V_beta_hat <- solve(t(X) %*% solve(Ve) %*% X)
    
    Q_tilde <- t(beta_hat) %*% solve(V_beta_hat) %*% beta_hat
    Q_df <- p - 1
    Q_p <- 1- pchisq(Q_tilde, df=Q_df)
    
    # significance test for H0: Beta_i = 0
    beta_hat <- as.vector(beta_hat)
    Z2 <- sapply(seq(beta_hat), function(X) beta_hat[X] / sqrt(V_beta_hat[X,X]))
    Z2_p_two_tail <- 1 - pnorm(abs(Z2))    
    

    results <- list(meta_table=data.frame(y, X),
         V_beta_hat=V_beta_hat,
         beta_hat=data.frame(beta_hat, z=Z2, p_two_tailed=Z2_p_two_tail),
         Q=c(Q=Q_tilde, df=Q_df, p=Q_p))
    class(results) <- c('fixed_effects_covariates_meta_analysis', 
                        'fixed_effects_meta_analysis')
    results
    
}


fixed_effects_covariates_meta_analysis(
    y = meta_data_covariates$y,
    X = as.matrix(meta_data_covariates[,c('x1', 'x2')]),
    sigma_sq = meta_data_covariates$sigma_sq)
                                                  
```


