The following implements some of the functions from Hunter and Schmidt (2004) page 95 onwards on "error of measurement and correction for attenuation.

# Correction based on true 
```{r}
correct_true_correlation <- function(rho_xy, rho_xx, rho_yy) {
    # calculate reliabilty based on population correlations
    # and reliabilities
    stopifnot(rho_xy >= -1 & rho_xy <= 1)
    stopifnot(rho_xx >= 0 & rho_xx <= 1)
    stopifnot(rho_yy >= 0 & rho_yy <= 1)
    stopifnot(rho_yy >= 0 & rho_yy <= 1)
    stopifnot(rho_xy <= (sqrt(rho_xx) * sqrt(rho_yy)))
    
    rho_xy / (sqrt(rho_xx) * sqrt(rho_yy))   
}

correct_true_correlation(.5, .8, .8)
```

# Uncorrected true correlation
Alternatively, the corrected true correlation may be known along with the reliability of observed x and observed y. From this, the uncorrected true correlation can be calculated.

```{r}
uncorrect_true_correlation <- function(rho_UT, rho_xx, rho_yy) {
     sqrt(rho_xx) * sqrt(rho_yy) * rho_UT
}

uncorrect_true_correlation(.5, .8, .8)
correct_true_correlation(.4, .8, .8) # restore
```

# Standard error of correlation
```{r}
standard_error_correlation <- function(r, n) {
    (1- r^2) / sqrt(n-1)
}

ci_correlation <- function(r, n, ci=.95) {
    z <- qnorm(1 - (1 - ci) /2)
    se_r <- standard_error_correlation(r, n)
    ci_diff <- z * se_r
    ci_r <- r + c(-1, 1) * ci_diff   
    names(ci_r) <- c('lower_ci', 'upper_ci')
    ci_r
}

ci_correlation_reliability_correct <- function(r, n, rho_xx, rho_yy, ci=.95) {
    ci_r <- ci_correlation(r, n, ci=ci)
    ci_rho <- correct_true_correlation(ci_r[1], rho_xx, rho_yy)
    ci_rho[2] <- correct_true_correlation(ci_r[2], rho_xx, rho_yy)
    names(ci_rho) <- c('lower_ci', 'upper_ci')
    ci_rho
}

corrected_error_variance <- function(rxy, n, rho_xx, rho_yy) {
    sigma_e_squared <- (1 - rxy^2)^2 / (n-1)
    reliability_denominator <- rho_xx * rho_yy
    sigma_e_squared / reliability_denominator
}

```


# Page 97 Example
The following code implements the example on page 97 of Hunter and Schmidt. 

```{r}
rho_TU <- .60
rho_xx <- .45
rho_yy <- .55
N <- 100
r_xy <- .20

# calculate true correlation
(rho_xy <- uncorrect_true_correlation(rho_TU, rho_xx, rho_yy))

# percentage reduction from corrected to uncorrected
round(((rho_TU - rho_xy) / rho_TU), 3)

# standard error of uncorrected correlation
standard_error_correlation(rho_xy, N)

# standard error of uncorrected observed correlation
standard_error_correlation(r_xy, N)

# ci uncorrected observed
round(ci_correlation(r_xy, N, .95), 2)

# ci corrected
round(ci_correlation_reliability_correct(r_xy, N, rho_xx, rho_yy, .95), 2)

# corrected observed correlation
(rc_xy <- correct_true_correlation(r_xy, rho_xx, rho_yy))

# corrected error variance
corrected_error_variance(r_xy, N, rho_xx, rho_yy)

# corrected standard error 
sqrt(corrected_error_variance(r_xy, N, rho_xx, rho_yy))
```
