# Fixed-effects meta analysis

The following document implements the fixed effects meta-analysis procedure outlined by Borenstein, Hedges, Higgins, and Rothstein (2009) pages 65 to 97.


# Generate sample data
```{r}
set.seed(1234)

sample_r_data <- function(true_r, n) {
    #true_r: true population correlations
    #n: vector of sample sizes for each study    
    rcor <- function(true_r, n) {
        x <- rnorm(n)
        y <- rnorm(n, true_r * x, sqrt(1 - true_r^2))
        cor(x, y)
    }
    
    # sample correlations
    Y <- sapply(n, function(X) rcor(true_r, n=X))
    
    # note that this is only approximate
    approx_variance <- function(r, n) (1 - r^2)^2 / (n - 1)
    
    # Variance of correlation 
    Vy <- approx_variance(Y, n)
    
    # Standard error of correlation
    SEy <- sqrt(Vy)
    
    data.frame(n, Y, Vy, SEy)
}
```


# Performing a fixed effect meta-analysis
```{r}
fixed_effect_meta_analysis <- function(Y, Vy, ci=.95) {
    # Y: vector of effect sizes (e.g., r correlations)
    # Vy: vector 
    # ci: scalar,  confidence interval: 0 < ci < 1
    stopifnot(length(Y) == length(Vy))
    
    # assign weight to each study
    W <- 1/ Vy
    
    # mean effect
    M <- sum(W * Y) /sum(W)
    
    # variance of summary effect
    Vm <- 1 / sum(W)
    
    # standard error of summary effect
    SEm <- sqrt(Vm)
    
    # Confidence interval for summary effect
    SEm_multiple <- abs(qnorm((1 - ci) / 2))
    LLm <- M - SEm * SEm_multiple
    ULm <- M + SEm * SEm_multiple
    
    # z-value for test of null hypothesis
    Z <- M / SEm
    
    # two tailed p-value
    p <- 2 * (1 - pnorm(abs(Z)))

    # return values
    list(studies=cbind(Y, Vy, W),
         M=M, Vm=Vm, SEm=SEm, ci=ci, ci_limits=c(LLm, ULm),
         test=c(Z=Z, p=p))
}
```


# Putting it all together
```{r}
meta1 <- sample_r_data(true_r = .3, n = seq(50, 250, 50))
fixed_effect_meta_analysis(meta1$Y, meta1$Vy)

meta2 <- sample_r_data(true_r = .1, n = rep(100, 10))
fixed_effect_meta_analysis(meta2$Y, meta2$Vy)
```


# Testing on examples from book
```{r}
# Table 14.2
fixed_effect_meta_analysis(
    Y = c(0.095, 0.277, 0.367, 0.664, 0.462, 0.185),
    Vy = c(0.033, 0.031, 0.050, 0.011, 0.043, 0.023))
```

The results are basically are approximately those provided on page 90 to 91. 
I used rounded values of $V_M$ and $Y$ from Table 14.2, and I used a more precise value for 95% confidence interval calculation.

